---
title: 'Class Assignment: MPG Tables and Charts'
author: "Tim Hogan"
date: "September 20, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)                ## loading the entire knitr package
library(ggplot2)              ## the only tidyverse library you need is ggplot2
library(esquisse)             ## use esquisse to setup the basic plots

library(kableExtra)
library(magrittr)
opts_chunk$set(echo = TRUE)
```

## Class assignment:

Using the MPG dataset, which is available with the ggplot2 library, produce a document that includes the following elements: headers, text, tables, and plots.  

Tables should include a comparison of city and highway mileage by the class of car and the best three car models for city and highway mileage for all the years in which data is available.  

Plot the data displaying as much of the data as you can. Put continuous variables on the axes and include a locally smoothed regression line to show the relationship with mileage.  Also make a box plot comparing city and highway MPG by class of car.  


## Table 1

This table was created by extracting the "class", "hwy", and "cty" columns, and using aggregate to summarize them by their means.

```{r, echo=FALSE}

data(mpg)     ## load the data

mpg$cyl <- as.factor(mpg$cyl)   ## convert discrete variables to factors so that they plot
mpg$drv <- as.factor(mpg$drv)   ## as different colors, not gradations of one color
mpg$cty <- as.double(mpg$cty)
```

```{r, echo=FALSE}                                

mpgclass <- mpg[, c("class","hwy","cty")]
mpgsummary <- aggregate(mpgclass[,2:3], list(mpg$class), mean)
kable(mpgsummary, booktabs=TRUE, digits = 2, col.names = c("Class", "Highway", "City"), align = "c") %>%
  kable_styling(font_size=18)
```


## Table 2

Once again, these tables were made by extracting the years, aggregating all models for all years, and then ordering hwy and cty to find the top ranks for all sets.

```{r, echo=FALSE}
mpg_99 <- mpg[mpg$year=="1999",] #Remove everything for desired year.
mpg_99summary <- aggregate(mpg_99[,c("cty","hwy")], list(mpg_99$model), mean) # Aggregate everything by model for desired variables.
mpg_99summary$cty <- round(mpg_99summary$cty,2) #Round values.
mpg_99summary$hwy <- round(mpg_99summary$hwy,2) #Round values.
mpg_99order.cty <- mpg_99summary[order(-mpg_99summary$cty),] #Order this column in a new data frame.
mpg_99order.hwy <- mpg_99summary[order(-mpg_99summary$hwy),] #Same with this. With this and the previous one, I chose not to remove the hwy column. It can easily be removed but I was interested in comparing them.

#This method was repeated for "2008"
mpg_08 <- mpg[mpg$year=="2008",]
mpg_08summary <- aggregate(mpg_08[,c("cty","hwy")], list(mpg_08$model), mean)
mpg_08summary$cty <- round(mpg_08summary$cty,2)
mpg_08summary$hwy <- round(mpg_08summary$hwy,2)
mpg_08order.cty <- mpg_08summary[order(-mpg_08summary$cty),]
mpg_08order.hwy <- mpg_08summary[order(-mpg_08summary$hwy),]

#This next thing will bind the columns for each desired set of groups and effects.
mpgtable <- cbind(mpg_99order.cty$Group.1[1:3],mpg_99order.cty$cty[1:3], 
                  mpg_99order.hwy$Group.1[1:3],mpg_99order.hwy$hwy[1:3],
                  mpg_08order.cty$Group.1[1:3],mpg_08order.cty$cty[1:3],
                  mpg_08order.hwy$Group.1[1:3],mpg_08order.hwy$hwy[1:3])

#Lastly, kable and subsequent functions will make something!
kable(mpgtable, digits = 3, booktabs=TRUE, col.names = c("Model", "Milage", "Model", "Milage","Model", "Milage", "Model", "Milage")) %>%
  add_header_above(c("City 1999"=2,"Highway 1999"=2,"City 2008"=2,"Highway 2008"=2)) %>%
  kable_styling(font_size=18)

```


## Plot 1

The codes for this table were created with esquisser().

```{r,echo=FALSE}

library(ggplot2)

#esquisser(mpg)


ggplot(mpg) +
 aes(x = displ, y = cty, colour = class) +
 geom_point(size = 1L) +
 scale_color_hue() +
 labs(x = "Engine Displacement", y = "City Mileage", title = "City MPG by Class of Car", subtitle = "1999 and 2008", color = "Car Class") +
 theme_minimal() +
 facet_wrap(vars(year))
```
 
## Plot 2
 
These ones were also made using esquisser(). 
 
```{r,echo=FALSE}
ggplot(mpg) +
 aes(x = displ, y = hwy, colour = class) +
 geom_point(size = 1L) +
 scale_color_hue() +
 labs(x = "Engine Displacement", y = "Highway Mileage", title = "Highway MPG by Class of Car", subtitle = "1999 and 2008", color = "Car Class") +
 theme_minimal() +
 facet_wrap(vars(year))
 
```

\newpage

## Boxplot 1

Once again, produced using esquisser().

```{r,echo=FALSE}
ggplot(mpg) +
 aes(x = class, y = cty) +
 geom_boxplot(fill = "#1f78b4") +
 labs(x = "Car Class", y = "City Mileage", title = "MPG in City by Car Class") +
 theme_linedraw(base_size=16)
```

\newpage

## Boxplot 2

esquisser() is my best friend!

```{r,echo=FALSE}
ggplot(mpg) +
 aes(x = class, y = hwy) +
 geom_boxplot(fill = "red") +
 labs(x = "Car Class", y = "Highway Mileage", title = "MPG in Highway by Car Class") +
 theme_linedraw(base_size=16)
```
